Log or not - Analysis of ECDC forecasts

### 2022-11-29 analysis/Analysis ECDC forecasts.Rmd
update plots

@@

  total_values <- 
    fulldata |> 
    filter(!is.na(prediction)) |>
    group_by(location, target_type, target_end_date) |>
    summarise(true_value = unique(true_value)) |>
    group_by(location, target_type) |>
    summarise(total = mean(true_value))

  plot_df <- scores |> 
    filter(model == "EuroCOVIDhub-ensemble", 
           horizon == 2) |>
    group_by(location, target_type, scale, type_and_scale) |>
    summarise(wis = summary_fct(interval_score), 
              .groups = "drop_last") |> 
    inner_join(total_values)

  plot_df |>
    group_by(target_type, scale) |>
    summarise(correlation = cor(wis, total)) |> 
    print()

  plot_df |>
    ggplot(aes(y = wis, x = total)) + 
    geom_point() + 
    facet_wrap(~ type_and_scale, scale = "free") + 
    theme_scoringutils() + 
    theme(legend.position = "none") + 
    scale_x_continuous(labels = scale_fn) +
    scale_y_continuous(labels = scale_fn) +
    labs(y = "REPLACE", x = "Total number of observed values")
}

@@

p_cor_scores <- df |>
  ggplot(aes(x = horizon, y = cor, 
             fill = `Target type`, color = `Target type`)) + 
  geom_histogram(stat = "identity", position = "dodge") +
  theme_scoringutils() + 
  coord_cartesian(ylim = c(0, 1)) +
  labs(y = "Avg. Spearman cor(scores)", x = "Forecast horizon")

@@

p_cor_skill <- correlation_rel_skill |>
  rename(`Target type` = target_type) |>
  ggplot(aes(x = horizon, y = cor, fill = `Target type`, color = `Target type` )) + 
  geom_histogram(stat = "identity", position = "dodge") +
  theme_scoringutils() + 
  coord_cartesian(ylim = c(0, 1)) +
  labs(y = "Cor(rel. skill)", x = "Forecast horizon")

@@

ggsave("output/figures/HUB-correlations.png", width = 7, height = 5.5)

@@

put_plot_together("EuroCOVIDhub-ensemble")
ggsave(filename = "output/figures/HUB-model-comparison-ensemble.png", width = 10, height = 7)

put_plot_together("epiforecasts-EpiNow2")
ggsave(filename = "output/figures/HUB-model-comparison-epinow.png", width = 10, height = 7)

put_plot_together("EuroCOVIDhub-baseline", locationname = "DE")
ggsave(filename = "output/figures/HUB-model-comparison-baseline.png", width = 10, height = 7)

@@

ranking_cases[[2]] + ranking_cases[[1]] + 
ranking_cases[[3]] + ranking_cases[[4]] + ranking_cases[[5]] + 
  ranking_deaths[[2]] + ranking_deaths[[1]] + 
ranking_deaths[[3]] + ranking_deaths[[4]] + ranking_deaths[[5]] +
  plot_annotation(tag_levels = "A") +
  plot_layout(design = layout, 
              guides = "collect") &
  theme(legend.position = "bottom")

>>>>

## =============================================================================
## FIGURE 1
# illustration-effect-log-score.R

## =============================================================================
## FIGURE 2
# Simulation poisson-mean-sd.R

## =============================================================================
## FIGURE 3
# example-log-first-Johannes.R

@@

  mean_values <- 
    fulldata |> 
    filter(!is.na(prediction)) |>
    group_by(location, target_type, target_end_date) |>
    summarise(true_value = unique(true_value)) |>
    group_by(location, target_type) |>
    summarise(mean_obs = mean(true_value))

  plot_df <- scores |> 
    filter(model == "EuroCOVIDhub-ensemble", 
           horizon == 2) |>
    group_by(location, target_type, scale, type_and_scale) |>
    summarise(wis = summary_fct(interval_score), 
              .groups = "drop_last") |> 
    inner_join(mean_values)

  plot_df |>
    group_by(target_type, scale) |>
    summarise(correlation = cor(wis, mean_obs)) |> 
    print()

  plot_df |>
    ggplot(aes(y = wis, x = mean_obs)) + 
    geom_point() + 
    facet_wrap(~ type_and_scale, scale = "free") + 
    theme_scoringutils() + 
    theme(legend.position = "none") + 
    scale_x_continuous(labels = scale_fn) +
    scale_y_continuous(labels = scale_fn) +
    labs(y = "REPLACE", x = "Mean observed value")
}

@@

p_cor_scores <- df |>
  ggplot(aes(x = horizon, y = cor, fill = `Target type`, color = `Target type` )) + 
  scale_fill_brewer(palette = "Set1", name = "Target type") + 
  scale_color_brewer(palette = "Set1", name = "Target type") +
  geom_histogram(stat = "identity", position = "dodge", alpha = 0.4) +
  theme_scoringutils() + 
  coord_cartesian(ylim = c(0, 1)) +
  labs(y = "Avg. rank cor(scores)", x = "Forecast horizon") + 
  theme(axis.title.y = element_text(size=8.5))

@@

p_cor_skill <- correlation_rel_skill |>
  rename(`Target type` = target_type) |>
  ggplot(aes(x = horizon, y = cor, fill = `Target type`, color = `Target type` )) + 
  scale_fill_brewer(palette = "Set1", name = "Target type") + 
  scale_color_brewer(palette = "Set1", name = "Target type") + 
  geom_histogram(stat = "identity", position = "dodge", alpha = 0.4) +
  theme_scoringutils() + 
  coord_cartesian(ylim = c(0, 1)) +
  labs(y = "Cor(rel. skill)", x = "Forecast horizon")

@@

ggsave("output/figures/HUB-correlations.png", width = 7, height = 7)

@@

put_plot_together("EuroCOVIDhub-ensemble")
ggsave(filename = "output/figures/HUB-model-comparison-ensemble.png", width = 10, height = 8.5)

put_plot_together("epiforecasts-EpiNow2")
ggsave(filename = "output/figures/HUB-model-comparison-epinow.png", width = 10, height = 8.5)

put_plot_together("EuroCOVIDhub-baseline", locationname = "DE")
ggsave(filename = "output/figures/HUB-model-comparison-baseline.png", width = 10, height = 8.5)

@@

ranking_cases[[2]] + ranking_cases[[1]] + 
ranking_cases[[3]] + ranking_cases[[4]] + ranking_cases[[5]] +
ranking_deaths[[2]] + ranking_deaths[[1]] + 
ranking_deaths[[3]] + ranking_deaths[[4]] + ranking_deaths[[5]] +
  plot_annotation(tag_levels = "A") +
  plot_layout(design = layout, 
              guides = "collect", 
              widths = c(1, 1, 1, 1, 2.5)) &
  theme(legend.position = "bottom")
